{\normalsize \textbf{Vagueness and imprecise imitation in signalling games}}

### 2016-05-31 paper/version_final/paper.tex
Adapted paper to BJPS template and bib style (using bibtex8 from now on for bib management).

\documentclass[fleqn,reqno,10pt]{article}

@@

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

@@

\title{Vagueness and imprecise imitation in signalling games}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

@@

@@

@@

@@

But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
could arise and be maintained in the first place. This is an apparent puzzle for functionalist
accounts that maintain that concepts and linguistic meanings evolved driven towards
efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signalling
situations, the existence of unclear borderline cases entails inefficiency of categorisation or
communication, or at least no advantage. The challenge is then to explain how vagueness
can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
without undermining the possibility of evolving, learning and communicating with a meaningful
language. A number of authors have consequently tried to explain why vagueness evolved as
something that is itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
latter line of thought.

We believe that vagueness in language and thought may have many reasons, not just one. We focus
here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
idea is that vagueness is, at least in part, due to imprecision in the perception of similar
stimuli and imprecision in the realisation of similar responses. On this view, vagueness in
language may be seen as a necessary sub-optimality due to limitations in another domain of
cognition, first and foremost perception. It could then be speculated that the whole system,
perception and language together, may be an almost optimal adaptation in a larger frame
of reference, e.g., if we take into account the metabolic costs for increased perceptual
accuracy. We will not engage in such speculation here. Rather we will explore the consequences
of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
of stimuli could come into play in a process of meaning evolution and how it could lead to
vague but by-and-large informative signal meaning. The formal model that this paper introduces
does exactly that. But it also does more. We find that confusability of stimuli can regularise
and systematise evolving meaning. This suggests a possibly advantageous side-effect that a
natural cause of vagueness may have and that might compensate some of the disadvantages that
vagueness may have for communicative efficiency \citep[c.f.][]{OConnor2013:The-Evolution-o}.

@@

% The view that vagueness is a natural concomitant of cognitive
% limitations of language users has been formalized in a number of ways,
% using evolutionary game theory and certain generalizations of Lewis'
% signaling games \citep{Lewis_1969:Convention}, so called
% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
% Our contribution is best seen in relation to these accounts, as it also
% relies on sim-max games. Let's introduce these first, and then zoom in
% on the problem of vagueness.

@@

Signalling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signalling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behaviour can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximising (short: sim-max) games are a variation of Lewis' games where the
receiver's actions are equated with the state space (one can think of the actions as choosing
states) and different states are allowed to be more or less similar to one another. While
Lewis' games treated communicative success as a matter of black and white, sim-max games allow
for shades of grey: the more similar the receiver's interpretation is to the actual state, the
better. Signalling games with utility-relevant similarities in the state space are fairly
standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor}, but have received
particular attention in a more philosophical context for reasons that will become clear
presently.

@@

\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of sim-max games with infinitely many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a quadratic loss function for utilities $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$ are remarkably systematic: the evolutionarily stable states are demonstrably so-called Voronoi languages.
Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the sender strategy partitions
the state space into convex categories, while the receiver's interpretations are the central
spots in each category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put, all points
in $X$ are connected via a straight line that lies entirely in $X$; $X$ has no gaps or
dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
language with two messages could have the sender use one message exclusively for all points in
the lower half of the unit interval and another for all points in the upper half; the
receiver's interpretations of messages are the central points, .25 and .75, in the respective
intervals.  See Figure~\ref{fig:Voronoi-language-example} for an illustration.

@@

        \caption{Example of a Voronoi language on $\States = [0,1]$. The pure sender strategy
          $\Spure$ uses one signal for the lower half, and another for the upper half of the
          unit interval. The pure receiver strategy $\Rpure$ selects the central elements in
          the respective intervals. See \citet{JagerMetzger2011:Voronoi-Languag} for further
          details.}
  \label{fig:Voronoi-language-example}
\end{figure}

It is intuitive to think that linguistic and conceptual categories are orderly in such a
manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
height of 2.2m, it would be difficult to defend not considering a person with a height
of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
project~\citep{CookKay2005:The-World-Color,WCS} collected colour naming data for 110 unwritten
languages of 45 language families.  For the great majority of these languages a pattern can be
observed: basic colour terms are by and large convex~\citep{Regier07,Jager2010}. It
is premature to argue that these observations can be extended to \emph{all} cases of
categorisation. However, for the cases where it does apply, the result of
\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
can impose this kind of orderly categories on a metric space without that being the ulterior
purpose of it all.%
\footnote{For the concrete case of colour categorisation, see also
  \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
Finally, these considerations are in line with a prominent
school of thought in comparative linguistics which also assumes that more abstract conceptual domains
(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
indefinite pronouns) are preferably carved up by the languages of the world in such a way
that meanings are connected regions on a so-called \emph{semantic map}
\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}.

There are at least two potential ways of interpreting the signalling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretise the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorisation
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

@@

\citet{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, a construction algorithm is available
that yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at a gradient transition between categories, by
weighing in the distance of different borderline cases to various
prototypical regions. This accounts for the first of the two
desiderata mentioned above, but still assumes that crisp prototype
regions must be given.

Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
met by evolving strategies in sim-max games. To illustrate what vague signalling would look
like, let us briefly consider a sim-max game with six equiprobable states and two messages and
what its equilibria would be like \citep[see][for a more thorough discussion of equilibria of
sim-max games]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that utilities
are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
three pairs of sender and receiver strategies. States are arranged according to their
similarity: the closer they are to each other, the more similar they are. The pair in
Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
signals is sub-optimal given the receiver's behaviour. Namely, if \mymessg{1} is interpreted as
\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
equilibrium in the sense that the speaker uses the same message for several states. Partial
pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
the two-message case. So, while partial pooling may entail inefficiency in some sense, and
while partial pooling can hamper the evolution of maximally efficient signal use
\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
close, so to speak. It shows smooth transitions across similar states at the boundaries of
categories and across acts around the most prototypical instances (as indicated by decreasing
thickness of arrows in Figure~\ref{fig:vagueUse}).

@@

The approach we take here is similar in spirit to that of
\citet{FrankeJager2010:Vagueness-Signa} and \citet{OConnor2013:The-Evolution-o}, but different
in relevant detail. A more in-depth comparison is deferred until
Section~\ref{sec:discussion}. Let us first motivate our approach here, and spell it out in more
detail in the following section.

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
or when they interact with functional pressure towards efficient communication, e.g., as
formalised in the replicator mutator dynamic
\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.

The emphasis of previous models that studied the effects of transmission infidelity on the
evolution of language has been on inductive biases and the systematicity, compression and
regularisation that they can introduce. Here, we would like to show that transmission noise of
a different kind can lead to regularisation as well and also give rise to vague meaning. We
find that shared perceptual biases that perturb the transmission of successful signalling
behaviour can regularise, facilitate and accelerate the evolution of meaning conventions, albeit
at the cost of vagueness. Concretely, we formalise the expected change in the behaviour of a
population of agents that try to imitate other agents' signalling behaviour. We assume, however,
that both realisation of behaviour, as well as observability of others' behaviour, is
systematically perturbed by noise. The resulting population-level dynamic generalises the
replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
evolutionary dynamic based on imitation
\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
confusability of stimuli does not undermine the possibility of evolving communicative signalling
behaviour. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
emergence of communicative signalling because it unifies and regularises evolutionary outcomes,
making it appear as if agents were applying inductive biases or generalising over partial
observations, when in truth this is the sole effect of confusion of similar stimuli.

@@

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signalling agents who imitate other agents' behaviour but
are liable to confuse states for one another. In the special limiting case where state
confusability vanishes, the process is just the well-known replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.

@@

\appendix

>>>>

\documentclass[12pt,english]{article}
\usepackage{etex}

@@

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
%\usepackage{todonotes}
\usepackage{subcaption}

@@

% paper size, margins, etc:
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\usepackage[titletoc,title]{appendix}

% spacing
\usepackage{setspace}
\doublespacing

% left justification:
\raggedright

% indentation
\setlength\parindent{12pt}

% section headings
\usepackage{sectsty}
\allsectionsfont{\normalsize\raggedright\centering}

% table of contents stuff
\setcounter{tocdepth}{2}
\usepackage[]{tocloft}
\addtocontents{toc}{\cftpagenumbersoff{section}}
\addtocontents{toc}{\cftpagenumbersoff{subsection}}
\renewcommand{\cftsecfont}{\itshape}
\renewcommand{\cftsubsecfont}{\itshape}
\makeatletter
\renewcommand{\@cftmaketoctitle}{}
\makeatother
\renewcommand{\cftdot}{}

% footnotes
\usepackage[]{footmisc}
\renewcommand\footnotelayout{\normalsize\doublespacing\raggedright}

% for bibliography
\usepackage{natbib}
\usepackage{url}
\makeatletter
\def\url@leostyle{%
  %\@ifundefined{selectfont}{
    \def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}
\makeatother
\urlstyle{leo}

\bibpunct
{(} %the opening bracket symbol
{)} %the closing bracket symbol
{,} %the punctuation between multiple citations
{a} %the letter `n' for numerical style, or `s' for numerical superscript style, any other letter for author-year
{} %the punctuation that comes between the author names and the year
{;} %the punctuation that comes between years or numbers when common author lists are suppressed

% headers
\pagestyle{plain}
% no page numbers:
%\pagestyle{empty}

% for author affiliation
\usepackage{authblk}

% misc
\usepackage{graphicx}
\usepackage[english]{babel}

% math packages
\usepackage{amsmath}
\usepackage{amsfonts}

% frames
%\usepackage{framed}

\makeatletter
\makeatother

% restart equation numbering after every section:
\numberwithin{equation}{section}

% FONTS:
%-------
% for times roman fonts:
\usepackage{txfonts}
% recommended if output will have accented characters:
\usepackage[T1]{fontenc}
% recommended if inputting accented directly from the keyboard:
%\usepackage[latin1]{inputenc}

% BJPS specific cite commands:
\newcommand{\citetbjps}[1]{\citeauthor{#1} ([\citeyear{#1}])}
\newcommand{\citetbjpspost}[2]{\citeauthor{#2} ([\citeyear{#2}], #1)}
\newcommand{\citealtbjps}[1]{\citeauthor{#1} [\citeyear{#1}]}
\newcommand{\citepbjps}[1]{(\citeauthor{#1} [\citeyear{#1}])}
\newcommand{\citepbjpspre}[2]{(#1~\citeauthor{#2} [\citeyear{#2}])}
\newcommand{\citeyearbjps}[1]{[\citeyear{#1}]}
\newcommand{\citeyearparbjps}[1]{([\citeyear{#1}])}

@@

\title{{\normalsize \textbf{Vagueness and imprecise imitation in signalling games}}}
\author{{\normalsize \textbf{Michael Franke}} \and {\normalsize \textbf{Jos\'e Pedro Correia}}}
\date{}

@@

%% no headers on the first page:
\thispagestyle{empty}

@@

\mbox{} \\

\tableofcontents

\mbox{} \\

@@

But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
could arise and be maintained in the first place. This is an apparent puzzle for functionalist
accounts that maintain that concepts and linguistic meanings evolved driven towards
efficiency. \citetbjps{Lipman2009:Why-is-Language} argues that, in common interest signalling
situations, the existence of unclear borderline cases entails inefficiency of categorisation or
communication, or at least no advantage. The challenge is then to explain how vagueness
can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
without undermining the possibility of evolving, learning and communicating with a meaningful
language. A number of authors have consequently tried to explain why vagueness evolved as
something that is itself useful
(e.g.~\citealtbjps{Jaegherde-Jaegher2003:A-Game-Theoreti}; \citealtbjps{Deemter2009:Utility-and-Lan}; \citealtbjps{BlumeBoard2013:Intentional-Vag}).
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citepbjpspre{e.g.}{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
learning strategies \citepbjpspre{e.g.}{OConnor2013:The-Evolution-o}. This paper contributes to the
latter line of thought.

We believe that vagueness in language and thought may have many reasons, not just one. We focus
here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
idea is that vagueness is, at least in part, due to imprecision in the perception of similar
stimuli and imprecision in the realisation of similar responses. On this view, vagueness in
language may be seen as a necessary sub-optimality due to limitations in another domain of
cognition, first and foremost perception. It could then be speculated that the whole system,
perception and language together, may be an almost optimal adaptation in a larger frame
of reference, e.g., if we take into account the metabolic costs for increased perceptual
accuracy. We will not engage in such speculation here. Rather we will explore the consequences
of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
of stimuli could come into play in a process of meaning evolution and how it could lead to
vague but by-and-large informative signal meaning. The formal model that this paper introduces
does exactly that. But it also does more. We find that confusability of stimuli can regularise
and systematise evolving meaning. This suggests a possibly advantageous side-effect that a
natural cause of vagueness may have and that might compensate some of the disadvantages that
vagueness may have for communicative efficiency \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.

@@

% The view that vagueness is a natural concomitant of cognitive
% limitations of language users has been formalized in a number of ways,
% using evolutionary game theory and certain generalizations of Lewis'
% signaling games \citepbjps{Lewis_1969:Convention}, so called
% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
% \citepbjps{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
% %\citepbjps{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
% Our contribution is best seen in relation to these accounts, as it also
% relies on sim-max games. Let's introduce these first, and then zoom in
% on the problem of vagueness.

@@

Signalling games, as introduced by \citetbjps{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signalling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behaviour can evolve is a highly interesting
topic that we are only beginning to fully understand
(e.g.~\citealtbjps{Warneryd1993:Cheap-Talk-Coor}; \citealtbjps{BlumeKim1993:Evolutionary-St}; \citealtbjps{Huttegger2007:Evolution-and-t}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{Barrett2009:The-Evolution-o}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}; \citealtbjps{Skyrms2010:Signals}).

Similarity-maximising (short: sim-max) games are a variation of Lewis' games where the
receiver's actions are equated with the state space (one can think of the actions as choosing
states) and different states are allowed to be more or less similar to one another. While
Lewis' games treated communicative success as a matter of black and white, sim-max games allow
for shades of grey: the more similar the receiver's interpretation is to the actual state, the
better. Signalling games with utility-relevant similarities in the state space are fairly
standard in economics
(e.g.~\citealtbjps{Spence1973:Job-market-sign}; \citealtbjps{CrawfordSobel1982:Strategic-Infor}), but have received
particular attention in a more philosophical context for reasons that will become clear
presently.

@@

\citetbjps{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of sim-max games with infinitely many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a quadratic loss function for utilities $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$ are remarkably systematic: the evolutionarily stable states are demonstrably so-called Voronoi languages.
Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the sender strategy partitions
the state space into convex categories, while the receiver's interpretations are the central
spots in each category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put, all points
in $X$ are connected via a straight line that lies entirely in $X$; $X$ has no gaps or
dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
language with two messages could have the sender use one message exclusively for all points in
the lower half of the unit interval and another for all points in the upper half; the
receiver's interpretations of messages are the central points, .25 and .75, in the respective
intervals.  See Figure~\ref{fig:Voronoi-language-example} for an illustration.

@@

        \caption{Example of a Voronoi language on $\States = [0,1]$. The pure sender strategy
          $\Spure$ uses one signal for the lower half, and another for the upper half of the
          unit interval. The pure receiver strategy $\Rpure$ selects the central elements in
          the respective intervals. See \citetbjps{JagerMetzger2011:Voronoi-Languag} for further
          details.}
  \label{fig:Voronoi-language-example}
\end{figure}

It is intuitive to think that linguistic and conceptual categories are orderly in such a
manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
height of 2.2m, it would be difficult to defend not considering a person with a height
of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
project~(\citealtbjps{CookKay2005:The-World-Color}; \citealtbjps{WCS}) collected colour naming data for 110 unwritten
languages of 45 language families.  For the great majority of these languages a pattern can be
observed: basic colour terms are by and large convex~(\citealtbjps{Regier07}; \citealtbjps{Jager2010}). It
is premature to argue that these observations can be extended to \emph{all} cases of
categorisation. However, for the cases where it does apply, the result of
\citetbjps{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
can impose this kind of orderly categories on a metric space without that being the ulterior
purpose of it all.%
\footnote{For the concrete case of colour categorisation, see also
  \citetbjps{JagerRooijvan-Rooij2007:Language-Struct} and \citetbjps{Correia2015}}
Finally, these considerations are in line with a prominent
school of thought in comparative linguistics which also assumes that more abstract conceptual domains
(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
indefinite pronouns) are preferably carved up by the languages of the world in such a way
that meanings are connected regions on a so-called \emph{semantic map}
(e.g.~\citealtbjps{Croft2003:Typology-and-Un}; \citealtbjps{Haspelmath2003:The-geometry-of}; \citealtbjps{LevinsonMeira2003:Natural-concept}).

There are at least two potential ways of interpreting the signalling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretise the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorisation
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further in
(\citealtbjps{Jager2007:The-Evolution-o}; \citealtbjps{JagerRooijvan-Rooij2007:Language-Struct}; \citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}),
among others.

@@

\citetbjps{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, a construction algorithm is available
that yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citetbjps{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at a gradient transition between categories, by
weighing in the distance of different borderline cases to various
prototypical regions. This accounts for the first of the two
desiderata mentioned above, but still assumes that crisp prototype
regions must be given.

Alternative approaches are taken by, for example, \citetbjps{FrankeJager2010:Vagueness-Signa} and
\citetbjps{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
met by evolving strategies in sim-max games. To illustrate what vague signalling would look
like, let us briefly consider a sim-max game with six equiprobable states and two messages and
what its equilibria would be like (see \citealtbjps{Jager2007:The-Evolution-o} and \citealtbjps{OConnor2013:The-Evolution-o} for a more thorough discussion of equilibria of
sim-max games). We assume that utilities
are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
three pairs of sender and receiver strategies. States are arranged according to their
similarity: the closer they are to each other, the more similar they are. The pair in
Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
signals is sub-optimal given the receiver's behaviour. Namely, if \mymessg{1} is interpreted as
\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
equilibrium in the sense that the speaker uses the same message for several states. Partial
pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
the two-message case. So, while partial pooling may entail inefficiency in some sense, and
while partial pooling can hamper the evolution of maximally efficient signal use
(e.g.~\citealtbjps{Hutteger:2007-Evol-Indicatives-Imperatives}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}),
this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
close, so to speak. It shows smooth transitions across similar states at the boundaries of
categories and across acts around the most prototypical instances (as indicated by decreasing
thickness of arrows in Figure~\ref{fig:vagueUse}).

@@

The approach we take here is similar in spirit to that of
\citetbjps{FrankeJager2010:Vagueness-Signa} and \citetbjps{OConnor2013:The-Evolution-o}, but different
in relevant detail. A more in-depth comparison is deferred until
Section~\ref{sec:discussion}. Let us first motivate our approach here, and spell it out in more
detail in the following section.

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
(\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
or when they interact with functional pressure towards efficient communication, e.g., as
formalised in the replicator mutator dynamic
(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}).

The emphasis of previous models that studied the effects of transmission infidelity on the
evolution of language has been on inductive biases and the systematicity, compression and
regularisation that they can introduce. Here, we would like to show that transmission noise of
a different kind can lead to regularisation as well and also give rise to vague meaning. We
find that shared perceptual biases that perturb the transmission of successful signalling
behaviour can regularise, facilitate and accelerate the evolution of meaning conventions, albeit
at the cost of vagueness. Concretely, we formalise the expected change in the behaviour of a
population of agents that try to imitate other agents' signalling behaviour. We assume, however,
that both realisation of behaviour, as well as observability of others' behaviour, is
systematically perturbed by noise. The resulting population-level dynamic generalises the
replicator dynamic \citepbjps{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
evolutionary dynamic based on imitation
(e.g.~\citealtbjps{Helbing1996:A-Stochastic-Be}; \citealtbjps{Schlag1998:Why-Imitate-and}). The inclusion of
confusability of stimuli does not undermine the possibility of evolving communicative signalling
behaviour. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
emergence of communicative signalling because it unifies and regularises evolutionary outcomes,
making it appear as if agents were applying inductive biases or generalising over partial
observations, when in truth this is the sole effect of confusion of similar stimuli.

@@

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signalling agents who imitate other agents' behaviour but
are liable to confuse states for one another. In the special limiting case where state
confusability vanishes, the process is just the well-known replicator dynamic
\citepbjps{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.

@@

\begin{appendices}
